The ChEMBL database as Linked Open Data

### 2012-04-03 article.tex
Completing the text of the triple structures

\hline
ChEMBL-RDF & Namespaces\\
act    & http://data.kasabi.com/dataset/chembl-rdf/activity/ \\
assay  & http://data.kasabi.com/dataset/chembl-rdf/assay/ \\
chembl & http://rdf.farmbio.uu.se/chembl/onto/\# \\
mol    & http://data.kasabi.com/dataset/chembl-rdf/molecule/ \\
res    & http://data.kasabi.com/dataset/chembl-rdf/resource/ \\
\hline
\end{tabular}
\end{table*}

@@

@@

@@

For each of the common resource classes a triple pattern was defined, following the
data available in the relational database. ... see Figure~\ref{f1}.

Assays .... anything

\begin{small}
\begin{verbatim}
assay:a17
 a chembl:Assay ;
 chembl:extractedFrom
  res:r11347 ;
 chembl:hasAssayType chembl:ADMET ;
 chembl:hasConfScore "7"^^xsd:int ;
 chembl:hasDescription
  "Inhibition of ... hydroxylase" ;
 chembl:hasTarget target:t100122 .
\end{verbatim}
\end{small}

Targets ... protein, other

\begin{small}
\begin{verbatim}
target:t1
 a chembl:Target ;
 rdfs:subClassOf pro:PR_000000001 ;
 rdfs:label "Glucoamylase" , 
  "Maltase-glucoamylase, intestinal" ;
 dc:identifier "uniprot:O43451" ,
   "3.2.1.3" ;
 dc:title "Maltase-glucoamylase" ;
 chembl:classL1 "Enzyme" ;
 chembl:hasDescription
  "Maltase-glucoamylase, intestinal" ;
 chembl:hasKeyword "Glycosidase" , 
  "Membrane" , "Sulfation" ;
 chembl:hasTaxonomy
  <http://bio2rdf.org/taxonomy:9606> ;
 chembl:organism "Homo sapiens" ;
 = <http://bio2rdf.org/uniprot:O43451> , 
   <http://bio2rdf.org/ec:3.2.1.3> .
\end{verbatim}
\end{small}

Activities ...

\begin{small}
\begin{verbatim}
act:a31863
 a       chembl:Activity ;
 cito:citesAsDataSource
  res:r6424 ;
 chembl:forMolecule mol:m180094 ;
 chembl:onAssay assay:a54505 ;
 chembl:relation ">" ;
 chembl:standardUnits
  "nM" ;
 chembl:standardValue
  "100000"^^xsd:float ;
 chembl:type "IC50" .
\end{verbatim}
\end{small}

\begin{tiny}
\begin{verbatim}
m41:inchikey
 a cheminf:CHEMINF_000059 ;
 cheminf:SIO_000300
   "LMCOMIDLRGMFCZ-RIPOXUOASA-N" .

mol:m41
 rdfs:label "CHEMBL406142" , 
   "Bis(3-[1 .... yl]-propionamide)" ;
 rdfs:subClassOf cheminf:CHEMINF_000000 ;
 cheminf:CHEMINF_000200
  m41:inchikey , m41:smiles , m41:inchi ;
 owl:equivalentClass 
  <http://rdf.openm .... 66+,67+,68+/m1/s1> ,
  <http://bio2rdf.org/chebi:100041> ,
  chemblid:CHEMBL406142 .
\end{verbatim}
\end{tiny}

@@

\begin{tiny}
\begin{verbatim}
journal:j6c706049c2e08871b7c46a6528065736
 a bibo:Journal ;
 dc:title "J. Med. Chem." .

res:r1
 a bibo:Article ;
 rdfs:seeAlso <http://bio2rdf.org/pubmed:14695813> ;
 dc:date "2004" ;
 dc:isPartOf
  journal:j6c706049c2e08871b7c46a6528065736 ;
 bibo:issue "1" ;
 bibo:pageEnd "9" ;
 bibo:pageStart "1" ;
 bibo:pmid "14695813" ;
 bibo:volume "47" .
\end{verbatim}
\end{tiny}

>>>>

\hline
ChEMBL-RDF & Namespace\\
\hline
chembl & http://rdf.farmbio.uu.se/chembl/onto/\# \\
\hline
ChEMBL-RDF & Prefixes\\
\hline
act    & http://data.kasabi.com/dataset/chembl-rdf/activity/ \\
assay  & http://data.kasabi.com/dataset/chembl-rdf/assay/ \\
mol    & http://data.kasabi.com/dataset/chembl-rdf/molecule/ \\
res    & http://data.kasabi.com/dataset/chembl-rdf/resource/ \\
\hline
\end{tabular}
\end{table*}

@@

An initial custom vocabulary has been defined, reflecting the concepts and links found in the
data structure of the ChEMBL database. This vocabulary has only partially been converted into
an OWL ontology, and is being phased out and replaced by common ontologies. This has not happened
for all data yet, and terms from the ChEMBL-RDF namespace (\textit{http://rdf.farmbio.uu.se/chembl/onto/\#},
prefix: \textit{chembl}) are still in use.

@@

For each of the common resource classes, a triple pattern was defined, following the
data available in the relational database. Figure~\ref{f1} shows how the various resource
classes are linked together. This section will show how data in the ChEMBL database
is exposed at an triple level.

The core concept in the ChEMBL database is that of the biological activity. This
is the type of information that the ChEMBL database extracts from literature.
Activities are at this moment still using an informal, custom vocabulary: the
type, fields and links to other resources are using the \textit{chembl}
namespace. While ChEMBL provides both the original activities as found in literature
and standardized values allowing comparison between studies, the triples only
make the latter available. The CiTO ontology is used to point to the paper from
which the data was extracted.

\begin{footnotesize}
\begin{verbatim}
act:a31863
 a       chembl:Activity ;
 cito:citesAsDataSource
  res:r6424 ;
 chembl:forMolecule mol:m180094 ;
 chembl:onAssay assay:a54505 ;
 chembl:relation ">" ;
 chembl:standardUnits
  "nM" ;
 chembl:standardValue
  "100000"^^xsd:float ;
 chembl:type "IC50" .
\end{verbatim}
\end{footnotesize}

The activities themselves are measured against assays, which make up a second important
resource type. Various assay types are found in the database: chembl:ADMET, chembl:Binding,
chembl:Functional, chembl:Property, and chembl:Unassigned. The assay information link
activities to a target (see below) and has a short description. The confidence information
assigned by the ChEMBL is exposed too, and was previously used to improve activity
prediction using Bayesian statistics~\cite{Willighagen2011}.

\begin{footnotesize}
\begin{verbatim}
assay:a17
 a chembl:Assay ;
 cito:citesAsDataSource
  res:r11347 ;
 chembl:hasAssayType chembl:ADMET ;
 chembl:hasConfScore "7"^^xsd:int ;
 chembl:hasDescription
  "Inhibition of ... hydroxylase" ;
 chembl:hasTarget target:t100122 .
\end{verbatim}
\end{footnotesize}

The assays measure activities against particular biological targets. The ChEMBL database
recognizes various types: pro:PR\_0000001, chembl:ADMET, chembl:CELL-LINE,
chembl:NUCLEIC-ACID, chembl:ORGANISM, chembl:SUBCELLULAR, chembl:TISSUE,
chembl:UNCHECKED, and chembl:UNKNOWN. The latter two are currently defined as
explicit types.

\begin{footnotesize}
\begin{verbatim}
target:t1
 a chembl:Target ;
 rdfs:subClassOf pro:PR_000000001 ;
 rdfs:label "Glucoamylase" , 
  "Maltase-glucoamylase, intestinal" ;
 dc:identifier "uniprot:O43451" ,
   "3.2.1.3" ;
 dc:title "Maltase-glucoamylase" ;
 chembl:classL1 "Enzyme" ;
 chembl:hasDescription
  "Maltase-glucoamylase, intestinal" ;
 chembl:hasKeyword "Glycosidase" , 
  "Membrane" , "Sulfation" ;
 chembl:hasTaxonomy
  <http://bio2rdf.org/taxonomy:9606> ;
 chembl:organism "Homo sapiens" .
\end{verbatim}
\end{footnotesize}

\begin{footnotesize}
\begin{verbatim}
m41:inchikey
 a cheminf:CHEMINF_000059 ;
 cheminf:SIO_000300
   "LMCOMIDLRGMFCZ-RIPOXUOASA-N" .

mol:m41
 rdfs:label "CHEMBL406142" , 
   "Bis(3-[1 .... yl]-propionamide)" ;
 rdfs:subClassOf cheminf:CHEMINF_000000 ;
 cheminf:CHEMINF_000200
  m41:inchikey , m41:smiles , m41:inchi ;
 owl:equivalentClass 
  <http://rdf.openm .... 66+,67+,68+/m1/s1> ,
  <http://bio2rdf.org/chebi:100041> ,
  chemblid:CHEMBL406142 .
\end{verbatim}
\end{footnotesize}

The molecular properties available from ChEMBL are now too exposed in triple format,
and like the InChI and InChIKey, these are provided using the CHEMINF ontology structure:

\begin{footnotesize}
\begin{verbatim}
mol:m1
 cheminf:CHEMINF_000200 m1:alogp .

m1:alogp
 a cheminf:CHEMINF_000305 ;
 cheminf:SIO_000300
  "3.344"^^xsd:double .
\end{verbatim}
\end{footnotesize}

@@

@@

\begin{footnotesize}
\begin{verbatim}
journal:j6c706049c2e08871b7c46a6528065736
 a bibo:Journal ;
 dc:title "J. Med. Chem." .

res:r1
 a bibo:Article ;
 rdfs:seeAlso <http://bio2rdf.org/pubmed:14695813> ;
 dc:date "2004" ;
 dc:isPartOf
  journal:j6c706049c2e08871b7c46a6528065736 ;
 bibo:issue "1" ;
 bibo:pageEnd "9" ;
 bibo:pageStart "1" ;
 bibo:pmid "14695813" ;
 bibo:volume "47" .
\end{verbatim}
\end{footnotesize}
